---
title: "Dining Locations Case Study"
author: "Noah Rubin (nar62)"
date: "October 26, 2015"
output: html_document
---
```{r}
# library from hadley wickham for web scraping inspired
# by beautiful soup (see: https://github.com/hadley/rvest)
library(xml2);
library(rvest);

namesURL = "https://living.sas.cornell.edu/dine/wheretoeat/";
labelsHTML = read_html(namesURL);
names(labelsHTML);
```
<p><b>
rvest provides a few main methods for parsing HTML: <br>
<p style="text-indent:50px;">
- html_node(".class #id node_type") --> extracts nodes that have class = class, id = id, type = node_type <br>
</p>
<p style="text-indent:50px;">
- e.x. html_node("strong span"), html_node(".reference a"), etc.
</p>
</b></p>
```{r}
linkNodes = labelsHTML %>%
  html_nodes("a");

links = linkNodes %>%
  html_attr("href");

labels = linkNodes %>%
  html_text();

# preview links and labels
head(links); head(labels);
```
<p><b>
Thus, we find that a lot of the links and labels are garbage, 
so it's time to clean them up. First, however, we will combine
them into a data frame for manipulation
</b></p>
```{r, fig.width=20}
diningLocations = as.data.frame(cbind(labels = labels, links = links));
# view data frame
diningLocations[44:80,]
```
<p><b>
As you can see from the frame above, cbind() combines the columns, and
as.data.frame coerces the combined columns into a data frame.  The columns
can be accessed by diningLocations\$links or diningLocations\$labels.  Now
I'm going to manually clean up the frame and only keep the relevant link:label
pairs.
</b></p>
```{r}
# only keep rows 44 to 78
diningLocations = diningLocations[45:78,];
# remove duplicates
diningLocations = diningLocations[!duplicated(diningLocations$labels),];
#remove unnecessary labels
exclude = c("convenience stores", "Cornell Dining Now");
diningLocations = diningLocations[-which(diningLocations$labels %in% exclude),];
# add domain to links
domain = "https://living.sas.cornell.edu";
diningLocations$links = paste(domain, diningLocations$links);
diningLocations$labels
# Looks like the duplicate function missed Jansen's Market
diningLocations = diningLocations[-which(diningLocations$labels == "Jansen’s Market"),];
```
<p><b>
Ok, for the cool part: getting latitude and longitude of each location.  As a first pass, I'm going to use the ggmap library, created by David Kahle and Hadley Wickham. Among other things, this lib provides a geocode method that returns the latitude and longitude of a given location. I'm also using a library called "purrr," produced by Hadley Wickham in the spirit of underscore.js.
</b></p>
```{r, results='hide', message=FALSE}
# ggplot 2 is dependency of ggmap
library(ggplot2);
# ggmap is library for map visualizations, and
# location informaiton and manipulations
library(ggmap);
library(purrr);

# locales stores the lat and longs given by geocode
locales = diningLocations$labels %>%
  # adding "Cornell University" for accuracy's sake
  paste(., " Cornell University", sep="") %>%
  # get lat and long
  map(geocode);
```
```{r}
head(locales);
```

<p><b>
Well, ggmap (as well as really any other library in R) sucks at getting coordinates for very specific and unpopular destinations like dining halls on west campus at Cornell University. Alas, all that is left is to change the information after the decimal points for lat and long, then we are done.
</b><p>
```{r}
# separate latitudes and longitudes for altering
latitudes = locales %>%
  map(function(position) position$lat) %>%
  unlist;
longitudes = locales %>%
  map(function(position) position$lon) %>%
  unlist;

# construct latitude decimal point alterations
lat_alts = c(.4473495, .4423221, .4534478, .4534452, .4535259, .453041, .4559407, .4535825, .4480805, .4457864, .4467172, .4484227, .4482542, .4466849, .4508134, .450099, .4467172, .4500624, .4443652, .446679, .4472154, .446891, .4478062, .4483435, .4488146, .4472589, .4468473, .4479873, .4439353);
# construct longitude decimal point alterations
long_alts = c(.4732484, .4872978, .4910123, .4910123, .4783553, .4841109, .4796581, .4783072, .4867747, .4854713, .4878249, .4832169, .4814839, .4863926, .4862022, .4836112, .4878249, .4814158, .4845526, .4857208, .484334, .4799113, .4815657, .4917249, .4916712, .4909148, .4916337, .4893484, .4897629);

# combine latitude and longitude corrections
latitudes = round(latitudes, 0) + lat_alts;
longitudes = round(longitudes, 0) - long_alts;
diningLocations$Latitude = latitudes;
diningLocations$Longitude = longitudes;

# fini! final data frame:
diningLocations[-2];
```
<p><b>
And that's it! If you want to write a file to a directory, follow these commands: <br>
<i> 1) setwd("path/to/directory"); </i> <br>
<i> 2) write.csv(data_frame, file="filename.csv"); </i> <br>
where data_frame is the dataframe object.  This of course assumes you want to write to .csv file, but there are plenty of write methods depending on the type of file you want to create.
</b></p>